Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-07-07 summary.tex
Update git repo

\begin{definition}[Subspace Detection Property]
Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy
points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the
solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of
$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from
different subspaces, then $X$ obeys the subspace detection property.
\end{definition}

@@

\begin{theorem}
\label{theorem5}
Let $P_n$ describe the edge probability matrix of the PABM with
$n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the
matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
positive and $K (K - 1) / 2$ most negative eigenvalues. Then
$\exists \lambda > 0$ and $N \in \mathbb{N}$ such that when $n > N$,
$\sqrt{n} \hat{V}_n$ obeys the subspace detection property with probability 1.
\end{theorem}

\begin{remark}
The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
from  \citeauthor{jmlr-v28-wang13} and the fact that the unscaled ASE of
$P_n$ consists of orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume
that the points in the embedding are all of unit length, and while we apply
this normalization in the simulated examples, it is not strictly necessary
for Theorem \ref{theorem5} due to orthogonality.
\end{remark}

@@

\begin{theorem}
\label{theorem6}
Under regularity and sparsity assumptions, given fixed $K$,

\begin{equation} \label{eq:thm6}
\max_{k, l \in \{1, ..., K\}} 
||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
\end{equation}
\end{theorem}

@@

We now provide a proof of Theorem~\ref{theorem5}. We first recall several
useful definitions for analyzing sparse subspace clustering.

\begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents
the inradius of the symmetric convex hull of $X$.
\end{definition}

\begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
\end{definition}

\begin{lemma}
\label{lemma2}
Let $\hat{V}$ be the eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues such that the rows of $\hat{V}$ are ordered by community, and let $\hat{V}^{(k)}$ be the rows of the $k^{th}$ community in $\hat{V}$ and $\hat{V}^{(-k)}$ be the rows of $\hat{V}$ with the $k^{th}$ community omitted. Denote $(\hat{v}_i^{(k)})^\top$ as the rows of $\hat{V}$, $\hat{V}_{-i}^{(k)}$ as $\hat{V}^{(k)}$ with the $i^{th}$ row omitted, and $\mathcal{S}^{(k)}$ as the subspace spanned by $V^{(k)}$. Let $V$, $V^{(k)}$, $V^{(-k)}$, and $v_i^{(k)}$ be the corresponding values for $P_n$.

Let $\nu_{i}^{(k)} = \max\limits_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2 \lambda} \eta^\top \eta$ subject to $||V_{-i}^{(k)} \eta||_\infty \leq 1$, and define the projected dual direction $w_{i}^{(k)}$ as $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$ normalized to length 1. Collect the projected dual directions into $W = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$.

Define the subspace incoherence:
$$\mu_n^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} ||W^{(k)} v||_\infty$$
Then $\forall k$, we have

\begin{equation} \label{eq:mu-conv}
\mu_n^{(k)} = 0
\end{equation}
\end{lemma}

\begin{proof} From Theorem \ref{theorem3} we have that the subspaces
  $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
  are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
  \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
  \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
  $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
    $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
    \in V^{(-k)}$ and hence $\mu_n^{(k)} = 0$ as desired.
%   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
% must be orthogonal to each row of \(V^{(-k)}\) and hence
%\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
\end{proof}

\begin{lemma}
\label{lemma3}
Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
$\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},

\begin{equation} \label{eq:delta-conv}
\delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
\end{equation}
\end{lemma}

\textbf{Proof of Theorem \ref{theorem5}}. The basis of this proof is
Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
subspace detection property holds if the noise is small enough and the
subspace inradius is greater than the subspace incoherence for each
community \(k\).

Let \(V_{n, -i}^{(k)}\) be \(V_n^{(k)}\) with the
\(i\)\textsuperscript{th} entry removed. Suppose that for each community
\(k\), there are enough vertices such that for each \(i\),
\(V_{n, -i}^{(k)}\) spans its corresponding subspace (Theorem
\ref{theorem2}). Then
\(r_n^{(k)} = \min\limits_i r(V_{n, -i}^{(k)}) > 0\). Thus by
(\ref{eq:mu-conv}), for each \(k\), \(r_n^{(k)} > \mu_n^{(k)} = 0\)
where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).

Let \(r_n = \min\limits_k r_n^{(k)}\). By (\ref{eq:delta-conv}),
\(\delta_n \stackrel{a.s.}{\to} 0\). Then as \(n\to \infty\),
\(\delta_n < \min\limits_k \frac{r_n (r_n^{(k)} - \mu_n^{(k)})}{2 + 7 r_n^{(k)}}\)
\(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
probability 1.

Thus the conditions for the subspace detection property from Theorem 6
from \citeauthor{jmlr-v28-wang13} are satisfied with probability 1 as
\(n \to \infty\).

\begin{remark}
Theorem 6 of \citeauthor{jmlr-v28-wang13} assume that each $\|v_n^{(i)}\| = 1$, 
which scales each $r_n^{(k)} \leq 1$. This is not strictly necessary for 
the proof of Theorem \ref{theorem5} since each $\mu_n^{(k)} = 0$, so as long 
as the $k^{th}$ community spans its subspace, $a r_n^{(k)} > 0 = \mu_n^{(k)}$ 
$\forall a > 0$.
\end{remark}

@@

\emph{Case \(k = l\)}. let $P^{(kk)} =
\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
decomposition of $P^{(kk)}$. We can then define
$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
singular value decomposition of \(A^{(kk)}\), and let
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
best rank-one approximation of $A^{(kk)}$. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = \frac{(\log n_k)^c}{\sqrt{n_k}}.$$
Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.

\emph{Case \(k \neq l\)}. \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
probabilities and edges between communities \(k\) and \(l\). Note that
\(P^{(kl)} = (P^{(lk)})^\top\).\\
By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). As
in the \(k = l\) case, we note that the singular value decomposition
\(P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top\) is one-dimensional
and \(\lambda^{(kl)} = \sigma_{kl} u^{(kl)}\). (We can also note that
the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
\(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
\(v^{(kl)} = u^{(lk)}\).)\\
Now consider the Hermitian dilation

\[M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}\]

which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. It can
be shown that the spectral decomposition of \(M^{(kl)}\) is

\[M^{(kl)} = 
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top\]

Thus treating \(M^{(kl)}\) as the edge probability matrix of a GRDPG, we
have latent positions in \(\mathbb{R}^2\) given by

\[\begin{bmatrix} 
  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
\end{bmatrix} = 
\begin{bmatrix} 
  \lambda^{(kl)} & \lambda^{(kl)} \\ 
  \lambda^{(lk)} & -\lambda^{(lk)} 
\end{bmatrix}\]

Now consider

\[\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}\]
Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where

\[E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}\]

and \(E\) is the \(n_k \times n_l\) matrix of independent noise (to
generate the Bernoulli entries in \(A^{(kl)}\).\\
Then \(\hat{M}^{(kl)}\) is an adjacency matrix drawn from \(M^{(kl)}\),
so its adjacency spectral embedding, given by

\[\begin{bmatrix} 
  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
\end{bmatrix}\]

>>>>

\begin{definition}[Subspace Detection Property]
  \label{def:subspace_detection}
Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy
points sampled from $K$ subspaces, i.e., $x_i = y_i + z_i$ where the
$y_i$ belongs to the union of $K$ subspaces and the $z_i$ are noise
vectors. Let $\lambda \geq 0$ be given and
let $C$ and $B$
be constructed from the
solutions of LASSO problems as described in Eq.~\eqref{eq:ssc} with
this given choice of $\lambda$. Then $X$
is said to satisfy the subspace detection property with weighting parameter $\lambda$ if each column of
$C$ has nonzero norm and $B_{ij} = 0$ whenever $y_i$ and $y_j$ are
from different subspaces.
\end{definition}

@@

\begin{theorem}
\label{theorem5}
Let $P_n$ describe the edge probability matrix of the PABM with
$n$ vertices, and let $A_n \sim \text{Bernoulli}(P_n)$.  Let $\hat{V}_n$ be the
matrix of eigenvectors of $A_n$ corresponding to the $K^2$ largest
eigenvalues in modulus. Then for all $\epsilon > 0$ 
there exists a choice of $\lambda > 0$ and $N \in \mathbb{N}$ such
that for all $n \geq N$, $\sqrt{n} \hat{V}_n$ obeys the subspace detection property with
probability at least $1 - \epsilon$.  
\end{theorem}

\begin{remark}
The proof of Theorem \ref{theorem5} is a direct consequence of Theorem 6
from  \citeauthor{jmlr-v28-wang13} and the fact that the columns of
$V_n$ correspond to orthogonal subspaces. \citeauthor{jmlr-v28-wang13} assume
that the points in the embedding are all of unit length, and while we apply
this normalization in the simulated examples, it is not strictly necessary
for Theorem \ref{theorem5} due to orthogonality.
\end{remark}

@@

\begin{theorem}
\label{theorem6}
Under regularity and sparsity assumptions, given fixed $K$, 
\begin{equation} \label{eq:thm6}
\max_{k, l \in \{1, ..., K\}} 
\|\hat{\lambda}^{(kl)} - \lambda^{(kl)}\|_{\infty} = 
O\bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
\end{equation}
with high probability.
\end{theorem}

@@

We now provide a proof of Theorem~\ref{theorem5}. Our proof is based
on verifying the sufficient conditions given in Theorem~6 of
\cite{jmlr-v28-wang13}
under which sparse subspace clustering based on solving the
optimization problem in Eq.~\eqref{eq:ssc} yields an affinity matrix
$B = |C| + |C^{\top}|$ satisfying the subspace detection property of
Definition~\ref{def:subspace_detection}. We first recall a few
definitions used in \cite{jmlr-v28-wang13}; for ease of exposition,
these definitions are stated using the notations of the current paper.

 % We first recall the
% notion of the inradius of a set of points. 
\begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
  \label{inradius}
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
Let $X$ be a $n \times d$ matrix with rows $x_1, x_2, \dots,
x_n$. We then define, with a slight abuse of notation, $r(X)$ as the
inradius of the convex hull formed by $\{\pm x_1, \pm x_2, \dots, \pm x_n\}$. 
\end{definition}

\begin{definition}[Subspace incoherence]
  \label{def:subspace_incoherence}
Let $\hat{V}$ be the eigenvectors of $A_n$
corresponding to the $K^2$ largest eigenvalues in modulus. Let
$\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
$\hat{V}$ corresponding to the $k^{th}$
community and let $\hat{V}^{(-k)}$ denote the matrix formed by
omitting the rows of $\hat{V}$
corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
$v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
$P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
rows of $V^{(k)}$.

Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
1,2,\dots,n_{k}$ as the solution of the following optimization problem
$$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
  \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
  \eta\|_\infty \leq 1$.}$$
Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
$\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
as
$$w_i^{(k)} =
\frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
and define the subspace incoherence for $\hat{V}^{(k)}$ by
$$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
\end{definition}

% Our proof is based on Theorem~6 in \cite{jmlr-v28-wang13} that provide
% sufficient conditions under which sparse subspace clustering based on
% solving the optimization problem in Eq.~\eqref{eq:ssc} yields an
% affinity matrix $B = |C| + |C^{\top}|$ satisfying the subspace
% detection property of Definition~\ref{def:subspace_detection}.  
% \begin{theorem}
%   Let $d \geq 1$ and let $\mathcal{S}_1, \mathcal{S}_2, \dots, \mathcal{S}_{L}$ be
%   subspaces of $\mathbb{R}^{d}$. 
% Let $Y = [y_1 \mid y_2 \mid \dots \mid y_n]^{\top}$ be a $n \times d$ matrix
% with rows $y_i \in \mathcal{S}_1 \cup \mathcal{S}_2 \cup
% \dots \mathcal{S}_{L}$. Let $x_i = y_i + z_i$ and suppose that there
% exists a finite $\delta \geq 0$ such that $\|z_i\| \leq \delta$ for
% all $i$.
% \end{theorem}

% Our proof leverages
% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
% guaranteeing the subspace detection property of a noisy

% Note that the rows
% of $\sqrt{n}V$ satisfy the subspace detection property in
% Definition~\ref{def:subspace_detection}. Now, by 
% Lemma~\ref{lem:technical}, there exists an orthogonal transformation
% $W$ such that the rows of $\sqrt{n}\hat{V}_n W$ are
% uniformly close to the rows of $\sqrt{n} V_n$.

% Our proof leverages
% the conditions given in Theorem~6 of \cite{jmlr-v28-wang13} for
% guaranteeing the subspace detection property of a noisy

% together with the uniform row-wise
% error bound between $\hat{V}$ and $V$ as given in
% Lemma~\ref{lem:technical}.

% We first recall the
% notion of the inradius of a set of points. 
% \begin{definition}[Inradius \cite{soltanolkotabi2012, jmlr-v28-wang13}]
% The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
% defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
% In addition, $r(X)$ for data matrix $X$ with rows $x_i^\top$ represents
% the inradius of the symmetric convex hull of $X$.
% \end{definition}

% The next

% \begin{definition}[Subspace incoherence property \cite{jmlr-v28-wang13}]
% \end{definition}

% \begin{lemma}
% \label{lemma2} % Let $\hat{V}$ be the eigenvectors of $A_n$
% % corresponding to the $K^2$ largest eigenvalues in modulus. Let
% % $\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
% % $\hat{V}$ corresponding to the $k^{th}$
% % community and let $\hat{V}^{(-k)}$ denote the matrix formed by
% % omitting the rows of $\hat{V}$
% % corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
% % the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
% % the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
% % $v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
% % $P_n$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
% % rows of $V^{(k)}$.

% % Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
% % 1,2,\dots,n_{k}$ as the solution of the following optimization problem
% % $$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
% %   \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
% %   \eta\|_\infty \leq 1$.}$$
% % Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
% % be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
% % $\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
% % as
% % $$w_i^{(k)} =
% % \frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
% % Now let $W^{(k)} = \begin{bmatrix} w_1^{(k)} & \cdots & w_{n_k}^{(k)} \end{bmatrix}^\top$
% % and define the subspace incoherence for $W^{(k)}$ by
% % $$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
% Recall the notations in Definition~\ref{def:subspace_incoherence}. 
% We then have
% \begin{equation} \label{eq:mu-conv}
% \mu^{(k)} = 0, \quad \text{for all $k = 1,2,\dots,K$.}
% \end{equation}
% \end{lemma}

% \begin{proof} From Theorem \ref{theorem3} we have that the subspaces
%   $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
%   are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
%   \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
%   \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
%   $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
%     $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
%     0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
%     \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ as desired.
% %   Therefore any vector projected to \(\mathcal{S}^{(k)}\)
% % must be orthogonal to each row of \(V^{(-k)}\) and hence
% %\(W^{(k)} v = 0\) for all $v \in \mathcal{S}^{(-k)}$, as desired.
% \end{proof}

% \begin{lemma}
% \label{lemma3}
% Let $(v_n^{(i)})^\top$ and $(\hat{v}_n^{(i)})^\top$ be the rows of $V_n$ and
% $\hat{V}_n$ respectively. By \citeauthor{rubindelanchy2017statistical},

% \begin{equation} \label{eq:delta-conv}
% \delta_n = \max_i \|\hat{v}_n^{(i)} - v_n^{(i)}\| \stackrel{a.s.}{\to} 0
% \end{equation}
% \end{lemma}

With the above definitions in place, we are now ready to state our
proof of Theorem~\ref{theorem5}.

\begin{proof}[Proof of Theorem~\ref{theorem5}]
For a given $k = 1,2\dots,K$, let $r^{(k)} = \min_{i}r(V_{-i}^{(k)})$ be inradius of the convex hull formed by
the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
\citeauthor{jmlr-v28-wang13} states that there exists a $\lambda > 0$
such that $\sqrt{n} \hat{V}$ satisfies
the subspace detection property in
Definition~\ref{def:subspace_detection} whenever
the following two conditions are satisfied
\begin{gather}
  \label{eq:cond1}
  \mu^{(k)} < r^{(k)} \quad \text{for all $k = 1,2,\dots,K$}, \\
  \label{eq:cond2}
  \max_{i} \|W \hat{v}_{i} - v_{i}\| \leq \min_{k} \frac{r_*(r^{(k)} -
    \mu^{(k)})}{2 + 7 r^{(k)}}.
\end{gather}
% Furthermore, we can also choose the weighting parameter $\lambda$ in
% the range
% $$\frac{1}{r_* - 2 \delta - \delta^2} \leq \lambda \leq \min_{k}
% \frac{r_k - \mu_k - 2 \delta_1}{\delta(1 + 
% In other words, $\sqrt{n}\hat{V}_n$ 
% small enough and the
% subspace inradius is greater than the subspace incoherence for each
% community \(k\).
We now verify that for sufficiently large n, Eq.~\eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
holds with high probability.
% satisfied with high probability. 
% Indeed, let
% The basis of this proof is Theorem 6 from \citeauthor{jmlr-v28-wang13}, which states that the
% subspace detection property holds if the noise in $\sqrt{n}\hat{V}$
% as an estimate for $\sqrt{n} V$ is small enough and that the inradius
% of the $\hat{V}^{(k)}$ are greater 
% small enough and the
% subspace inradius is greater than the subspace incoherence for each
% community \(k\).
 % , let \(v_{-i}^{(k)}\) be the matrix \(V^{(k)}\) with the
% \(i\)\textsuperscript{th} row removed. Then

{\bf Verifying Eq.~\eqref{eq:cond1}}. If $n$ is sufficiently large then
there are enough vertices in each community $k$ so that
$\mathrm{span}(V_{-i}^{(k)}) = \mathcal{S}^{(k)}$ for all $i$ and hence
% spans its corresponding subspace (Theorem
% \ref{theorem2}). 
% We therefore have, for sufficiently large $n$, that 
\(r^{(k)} = \min_{i} r(V_{-i}^{(k)}) > 0\) for
all $k = 1,2,\dots,K$. % Eq.~\eqref{eq:cond1} then follows from the fact that $\mu^{(k)} = 0$ for all $k = 1,2,\dots,K.$
% Indeed, from 
Next, by Theorem \ref{theorem3} we have that the subspaces
  $\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
  are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
  \mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
  \ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
  $\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
    $z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
    0$ for all $v \in V^{(-k)}$. Since $z$ is arbitrary, this implies $\|W^{(k)} v\|_{\infty} = 0$ for all $v
    \in V^{(-k)}$ and hence $\mu^{(k)} = 0$ for all $k
    =1,2,\dots,K$. Therefore $\mu^{(k)} < r^{(k)}$ for all $k =
    1=2,\dots,K$ as desired.

% Lemma~\ref{lemma2} then implies, with high
% Combining the above two facts, we have, for sufficiently large, that $n$ \(r^{(k)} > \mu_n^{(k)} = 0\) for all $k = 1,2\dots,K$.
% where \(\mu_n^{(k)} = \mu(\hat{V}_n^{(k)})\) and \(n\) is large enough
% such that \(\min\limits_{k, i} \text{rank}(V_{n, -i}^{(k)}) = K\).

{\bf Verifying Eq.~\eqref{eq:cond2}}.
Let $\delta = \max_{i} \sqrt{n} \|W
\hat{v}_{i} - v_{i}\|$. Then from Lemma~\ref{lem:technical}, we have
\(\delta \stackrel{a.s.}{\to} 0\) and hence
$$\delta < \min_{k} \frac{r_* (r^{(k)} - \mu^{(k)})}{2 + 7 r^{(k)}}$$
asymptotically almost surely. 
% \(= \min\limits_k \frac{r_n r_n^{(k)}}{2 + 7 r_n^{(k)}}\) with
% probability 1.

In summary $\sqrt{n} \hat{V}$ satisfies the subspace detection property
with probability converging to $1$ as \(n \to \infty\).
\end{proof}

\begin{remark}
Theorem 6 of \citeauthor{jmlr-v28-wang13} assumes that each row $v_i$
of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. This
assumption has the effect of scaling the $r^{(k)}$ so that $r^{(k)}
\leq 1$ for all $k = 1,2,\dots,K$. We emphasize that this assumption
has no effect on the proof of Theorem \ref{theorem5}. Indeed,
since $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$
spans the subspace $S^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ 
for any scalar $a > 0$. 
\end{remark}

@@

\emph{Case \(k = l\)}. Let $P^{(kk)} =
\sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ be the singular value
decomposition of $P^{(kk)}$. We can then define
$\tilde{\lambda}^{(kk)} = \sigma_{kk} u^{(kk)}$. 
Now let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the
singular value decomposition of \(A^{(kk)}\), and let
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be the
best rank-one approximation of $A^{(kk)}$. Define
\(\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}\). Then
\(\hat{\lambda}^{(kk)}\) is the adjacency spectral embedding approximation of \(\lambda^{(kk)}\)
and by Theorem 5 of \citeauthor{rubindelanchy2017statistical}, we have
%adjacency spectral embedding \(\hat{\lambda}^{(kk)}\) approximates
$$\|\hat{\lambda}^{(kk)} - \lambda^{(kk)}\|_{\infty} = O\Bigl(\frac{(\log n_k)^c}{\sqrt{n_k}}\Bigr)$$
with high probability. 
Here $\|\cdot\|_{\infty}$ denote the $\ell_{\infty}$ norm of a vector.

\emph{Case \(k \neq l\)}. % \(P^{(kl)}\) and \(A^{(kl)}\) represent edge
% probabilities and edges between communities \(k\) and \(l\). Note that
% \(P^{(kl)} = (P^{(lk)})^\top\).\\
% By definition, \(P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top\). 
Let \(P^{(k \ell)} = \sigma_{k \ell}^2 u^{(k \ell)} (v^{(k \ell)})^\top\) and 
\(P^{(\ell k)} = \sigma_{kl}^2 u^{(\ell k)} (v^{(\ell k)})^\top\)  be the singular
value decompositions and note that \(\sigma_{k \ell} = \sigma_{\ell
  k}\), \(u^{(k \ell)} = v^{(\ell k)}\), and
\(v^{(k \ell)} = u^{(\ell k)}\). 
Now define \(\lambda^{(k \ell)} = \sigma_{k \ell} u^{(k \ell)}\) and
\(\lambda^{(\ell k)} = \sigma_{k \ell}
v^{(k \ell)}\).

% (We can also note that
% the SVD of \(P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top\), i.e.,
% \(\sigma_{kl} = \sigma_{lk}\), \(u^{(kl)} = v^{(lk)}\), and
% \(v^{(kl)} = u^{(lk)}\).)\\
Next consider the Hermitian dilation

\[M^{(k \ell)} = 2 \begin{bmatrix} 0 & P^{(k \ell)} \\ P^{(\ell k)} & 0 \end{bmatrix}\]

which is a symmetric \((n_k + n_l) \times (n_k + n_l)\) matrix. The
eigendecomposition of \(M^{(kl)}\) is then

\[M^{(kl)} = 
\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix} \times 
\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
\begin{bmatrix} u^{(k \ell)} & -u^{(k \ell)} \\ v^{(k \ell)} & v^{(k \ell)} \end{bmatrix}^\top\]
Thus treating \(M^{(k \ell)}\) as the edge probability matrix of a GRDPG, we
have latent positions in \(\mathbb{R}^2\) given by the $(n_k + n_{\ell}) \times 2$ matrix
\[\Lambda^{K \ell} = \begin{bmatrix} 
  \sigma_{k \ell} u^{(k \ell)} & \sigma_{k \ell} u^{(k \ell)} \\ 
  \sigma_{k \ell} v^{(k \ell)} & -\sigma_{k \ell} v^{(k \ell)} 
\end{bmatrix} = 
\begin{bmatrix} 
  \lambda^{(k \ell)} & \lambda^{(k \ell)} \\ 
  \lambda^{( \ell k)} & -\lambda^{( \ell k)} 
\end{bmatrix}.\]
Now consider
\[\hat{M}^{(k \ell)} = \begin{bmatrix} 0 & A^{(k \ell)} \\ A^{(\ell k)} & 0 \end{bmatrix}\]
% Then \(\hat{M}^{(kl)} = M^{(kl)} + E'\) where
